---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r message=FALSE}
rm(list = ls())
library("dplyr")
library("tidyr")
library("ggplot2")
data <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
1. Histogram of the total number of steps taken each day
```{r}
step_sum <- data %>% group_by(date) %>% summarise(ssum=sum(steps,na.rm = TRUE))
ggplot(step_sum)+
    geom_histogram(aes(ssum),binwidth = 2000)+
    labs(x="Totoal number of steps per day",y="Frequency")
```

2. The mean and median of the total number of steps taken per day
```{r}
step_mean <- data %>% group_by(date) %>% summarise(smean=mean(steps,na.rm = TRUE))
step_mean
step_median <- data %>% group_by(date) %>% summarise(smedian=median(steps,na.rm = TRUE))
step_median
```

## What is the average daily activity pattern?
1. Time series plot of the number of steps taken averaged across all days
```{r}
step_ave <- data %>% 
    group_by(interval) %>%
    summarise(avg=mean(steps,na.rm = TRUE))
plot(step_ave$avg,xlab = "Interval",ylab = "Average number of steps",type="l")
```

2. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
step_ave[which.max(step_ave$avg),]
```

## Imputing missing values
1. Total number of rows with NAs
```{r}
sum(is.na(data$steps))
```
2. The missing data are replaced by the mean values of the steps for all days 

3. Create a new dataset data_new that is equal to the original dataset but with missing data filled in
```{r}
data_new <- replace_na(data,list(steps=mean(data$steps,na.rm = TRUE)))
```

4. Plot histogram of the total number of steps taken each day after missing values are imputed
```{r}
step_sum_new <- data_new %>% group_by(date) %>% summarise(ssum=sum(steps,na.rm = TRUE))
ggplot(step_sum_new)+
    geom_histogram(aes(ssum),binwidth = 2000)+
    labs(x="Totoal number of steps per day",y="Frequency")
```  

The mean and median of the total number of steps taken per day
```{r}
step_mean_new <- data_new %>% group_by(date) %>% summarise(smean=mean(steps,na.rm = TRUE))
step_mean_new
step_median_new <- data_new %>% group_by(date) %>% summarise(smedian=median(steps,na.rm = TRUE))
step_median_new
```
Imputing missing data increases the total number of daily steps

## Are there differences in activity patterns between weekdays and weekends?
Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
data_new <- mutate(data_new,day=ifelse(weekdays(as.Date(data_new$date)) %in% c("Saturday","Sunday"), "Weekend", "Weekday"))

step_Weekdays <- data_new %>% 
    filter(day=="Weekday") %>%
    group_by(interval) %>%
    summarise(avg=mean(steps,na.rm = TRUE))

step_weekends <- data_new %>% 
    filter(day=="Weekend") %>%
    group_by(interval) %>%
    summarise(avg=mean(steps,na.rm = TRUE))

par(mfrow=c(2,1))
plot(step_Weekdays$avg,xlab = "Interval",ylab = "Ave. steps",type="l",main="Weekday")
plot(step_weekends$avg,xlab = "Interval",ylab = "Ave. steps",type="l",main="Weekend")
```